5'CutSiteAnalysisDataGeneration

### 2023-07-31 R/CutSiteCorrelationDataGeneration.rmd
Improved cut site correlation analysis
- Included option to use normal mismatches by reads files with minimal filtering
- Included analysis of reads with shuffled 3' cut site distances as a negative control (i.e., a sanity check).

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysisDataGeneration = function(lesion, repetition, timepoint, cellType, inputDirectory, outputDirectory,
                                                    includedMismatches = c("CC>TT")) {
  mismatchesByRead = fread(generateFilePath(
    inputDirectory, dataTypeStrings$mismatchesByReadRelationFormatted , ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))
  fwrite(getPairedCutSiteDistances(mismatchesByRead),  sep = '\t', file = generateFilePath(
    outputDirectory, dataTypeStrings$pairedCutSiteDistances , ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))

@@

```{r generate_cut_site_correlation_data}
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory)
```

>>>>

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysisDataGeneration = function(lesion, repetition, timepoint, cellType, inputDirectory, outputDirectory,
                                                    includedMismatches = c("CC>TT"), filteredAndFormattedInput = TRUE,
                                                    filePathIncludedMismatches = NA) {

  if (length(filePathIncludedMismatches) > 0 && is.na(filePathIncludedMismatches)) {
    filePathIncludedMismatches = includedMismatches
  }

  if (filteredAndFormattedInput) {
    inputType = dataTypeStrings$mismatchesByReadRelationFormatted
  } else {
    inputType = dataTypeStrings$mismatchesByRead
  }

  mismatchesByRead = fread(generateFilePath(
    inputDirectory, inputType, ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = filePathIncludedMismatches
  ))

  if (filteredAndFormattedInput) {
    additionalInformation = character()
  } else {
    additionalInformation = "minimal_filtering"
  }

  if (length(filePathIncludedMismatches) != length(includedMismatches)) {
    mismatchesByRead = mismatchesByRead[V5 %in% includedMismatches]
    mismatchesByRead[,V4:=as.numeric(V4)]
  }

  fwrite(getPairedCutSiteDistances(mismatchesByRead, filteredAndFormattedInput), sep = '\t', file = generateFilePath(
    outputDirectory, dataTypeStrings$pairedCutSiteDistances , ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches,
    additionalInformation = additionalInformation
  ))

@@

```{r generate_cut_site_correlation_data}
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory)
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory,
                                         includedMismatches = "C>T")
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory,
                                         includedMismatches = "C>T", filteredAndFormattedInput = FALSE,
                                         filePathIncludedMismatches = character())
cutSiteCorrelationAnalysisDataGeneration("CPD", "all_reps", "1h", "NHF1",
                                         NHF1MismatchesByReadDirectory, NHF1CutSiteCorrelationDirectory,
                                         includedMismatches = "CC>TT", filteredAndFormattedInput = FALSE)
```

### 2023-07-31 R/CutSiteCorrelationFigureGeneration.rmd
Improved cut site correlation analysis
- Included option to use normal mismatches by reads files with minimal filtering
- Included analysis of reads with shuffled 3' cut site distances as a negative control (i.e., a sanity check).

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysisFigureGeneration = function(lesion, repetition, timepoint, cellType, directory,
                                                    includedMismatches = c("CC>TT")) {
  pairedCutSiteDistances = fread(generateFilePath(
    directory, dataTypeStrings$pairedCutSiteDistances , ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches
  ))

  print(getCorrelationResults(pairedCutSiteDistances))
  plotPairedCutSiteDistances(pairedCutSiteDistances,
                             title = paste(cellType, lesion, timepoint, includedMismatches, "Cut Site Correlation"))

@@

```{r generate_cut_site_correlation_figures}
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory)
```

>>>>

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteCorrelationAnalysisFigureGeneration = function(lesion, repetition, timepoint, cellType, directory,
                                                      includedMismatches = c("CC>TT"), filteredAndFormattedInput = TRUE,
                                                      shuffle = FALSE) {

  if (filteredAndFormattedInput) {
    additionalInformation = character()
  } else {
    additionalInformation = "minimal_filtering"
  }

  pairedCutSiteDistances = fread(generateFilePath(
    directory, dataTypeStrings$pairedCutSiteDistances , ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, filtering = filtering, includedMismatches = includedMismatches,
    additionalInformation = additionalInformation
  ))

  if (shuffle) {
    pairedCutSiteDistances[,Three_Prime_Cut_Site_Distance :=
                             sample(Three_Prime_Cut_Site_Distance, nrow(pairedCutSiteDistances))]
  }

  correlationResults = getCorrelationResults(pairedCutSiteDistances)
  print(correlationResults)
  print(paste("True p-value:",correlationResults$p.value))

  plotPairedCutSiteDistances(pairedCutSiteDistances,
                             title = paste(cellType, lesion, timepoint, includedMismatches,
                                           "Cut Site Correlation", additionalInformation))

@@

```{r generate_cut_site_correlation_figures}
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory)
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory,
                                           includedMismatches = "C>T")
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory,
                                           filteredAndFormattedInput = FALSE)
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory,
                                           includedMismatches = "C>T", filteredAndFormattedInput = FALSE)
```

```{r shuffled_for_sanity_check}
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory,
                                           shuffle = TRUE)
cutSiteCorrelationAnalysisFigureGeneration("CPD", "all_reps", "1h", "NHF1", NHF1CutSiteCorrelationDirectory,
                                           includedMismatches = "CC>TT", filteredAndFormattedInput = FALSE,
                                           shuffle = TRUE)
```
